---
title: "Exploratory Data Analysis script"
output: html_document
author: Antonio Flores
---


This Quarto file loads the cleaned data and does some exploring.

# Setup
Necessary libraries for this script.
```{r}
#| output: false
library(here) 
library(dplyr)
library(skimr)
library(ggplot2)
library(scales)
library(tidyr)
library(forcats)
library(corrplot)
library(corrplot)
library(knitr)
library(kableExtra)
```


Loading the data.

```{r}
#If this doesn't initially work, try restarting R and then proceeding
data_location <- here::here("data","processed-data","processeddata1.rds")
mydata <- readRDS(data_location)
```


# Data exploration through tables

Creating a summary table.

```{r}
summary_df = skimr::skim(mydata)
print(summary_df)

# save to file
summarytable_file = here("results", "tables", "summarytable1.rds")
saveRDS(summary_df, file = summarytable_file)
```




```{r}
data_location <- here::here("data","processed-data","processeddata1.rds")
mydata1 <- readRDS(data_location)

mydata1 = mydata1 %>% 
  select(age, job, marital, education, default, balance, housing, loan, contact, day, month, y_termSubscribed)

str_summary <- capture.output(str(mydata1))

describetbl1 = kable(str_summary, caption = "Summary of predictors") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), full_width = F)

summarytable_file = here("results", "tables", "describetbl1.rds")
saveRDS(describetbl1, file = summarytable_file)


describetbl1
```




























# Data exploration through figures

Histogram plots for some continuous outcomes. 

Age first.

```{r}
p1 <- mydata %>% ggplot(aes(x=age)) + geom_histogram()
plot(p1)
figure_file = here("results", "figures","age-distribution.png")
ggsave(filename = figure_file, plot=p1) 
```

Next bank account balance.

```{r}
p2 <- mydata %>% ggplot(aes(x=balance)) + geom_histogram() + scale_x_log10(labels =label_dollar())
plot(p2)
figure_file = here("results", "figures","balance-distribution.png")
ggsave(filename = figure_file, plot=p2) 
```

Now balance as function of age

```{r}
p3 <- mydata %>% ggplot(aes(x=age, y=balance)) + geom_point() + geom_smooth(method='lm')
plot(p3)
figure_file = here("results", "figures", "age-balance-scatter.png")
ggsave(filename = figure_file, plot=p3) 
```

Now balance as function of age, stratified by the marital variable. 

```{r}
p4 <- mydata %>% ggplot(aes(x=age, y=balance, color = marital)) + geom_point() + geom_smooth(method='lm')
plot(p4)
figure_file = here("results", "figures", "age-balance-stratified.png")
ggsave(filename = figure_file, plot=p4) 
```


Next, exploring balance across different job types. 

```{r}
p5 = mydata |>
  ggplot(aes(balance, job))+
  geom_boxplot()+
  scale_x_log10(labels =label_dollar())
p5
figure_file = here("results", "figures", "job-balance-scatter.png")
ggsave(filename = figure_file, plot=p5) 

```

Now, some barplots examining job types, education level, and marital status.

```{r}
p6 = mydata |>
  ggplot(aes(x=fct_infreq(job)))+
  geom_bar() +
  coord_flip()

p7 = mydata |>
  ggplot(aes(education))+
  geom_bar()

p8 = mydata |>
  ggplot(aes(marital))+
  geom_bar()


p6
p7
p8

figure_file = here("results", "figures", "job-barchart.png")
ggsave(filename = figure_file, plot=p6) 
figure_file = here("results", "figures", "education-barchart.png")
ggsave(filename = figure_file, plot=p7) 
figure_file = here("results", "figures", "marital-barchart.png")
ggsave(filename = figure_file, plot=p8) 
```



Lets see how the age demographic differs between the population that subscribed to a term (y_termsubscribed=yes) vs the general population.

```{r}
p9 = mydata |>
  filter(y_termSubscribed == 1) |>
  ggplot(aes(age))+
  geom_bar()+
  labs(title = "yes")

p10 = mydata |>
  ggplot(aes(age))+
  geom_bar()+
  labs(title = "yes/no")

p9
p10

figure_file = here("results", "figures", "yes-age.png")
ggsave(filename = figure_file, plot=p9) 

figure_file = here("results", "figures", "yes-no-age.png")
ggsave(filename = figure_file, plot=p10) 



#Comparing how age data compares from those who subscribed to a term deposit and those
#that did not
ggplot(data=mydata)+
  geom_bar(mapping =aes(x=age))+
  facet_wrap(~y_termSubscribed, nrow=2)


```


Now, lets assess the highest job categories that are labeled as subscribing to a term.

```{r}
p11 = mydata |>
  filter(y_termSubscribed == 1) |>
  ggplot(aes(x=fct_infreq(job)))+
  geom_bar()+
  coord_flip()+
  labs(title = "Term Subscribed = yes")

p11
```









Taking a quick look at some of the binary variables in the dataset
```{r}
ggplot(data=mydata)+
  geom_bar(mapping =aes(x=default))
ggplot(data=mydata)+
  geom_bar(mapping =aes(x=housing))
ggplot(data=mydata)+
  geom_bar(mapping =aes(x=loan))
```



Taking a look at frequency counts across different variables of interest
```{r}
hist(mydata$age)
hist(mydata$balance)
```


Examining which days are most common for recording a positive outcome (subscribing to 
a term deposit). Looks like later in the month has higher frequencies.

```{r}
p12= mydata |>
  filter(y_termSubscribed == 1) |>
  ggplot(aes(day))+
  geom_bar()+
  labs(title = "yes")

p12
figure_file = here("results", "figures", "days-yes.png")
ggsave(filename = figure_file, plot=p12) 
```



We've analyzed which job types had the MOST positive outcomes, but now we want to 
compare which job type had the greatest PERCENT of positive outcomes.

```{r}
posjob = mydata %>% 
  filter(y_termSubscribed==1) %>% 
  count(y_termSubscribed, job, sort=TRUE)

jobtable = mydata %>% count(job)

pctjob = posjob %>% 
  left_join(jobtable, by ="job")


colnames(pctjob)[3] <- "numofYES"
colnames(pctjob)[4] <- "numofBOTH"

```

```{r}
p13 = ggplot(pctjob, aes(x=reorder(job, numofYES), y=numofYES, fill=numofYES))+
  geom_bar(stat="identity")+
  coord_flip()
  
p14 = ggplot(pctjob, aes(x=reorder(job, (numofYES/numofBOTH)), y=(numofYES/numofBOTH), fill=numofYES))+
  geom_bar(stat="identity")+
  coord_flip()

figure_file = here("results", "figures", "jobs-count-yes.png")
ggsave(filename = figure_file, plot=p13) 
figure_file = here("results", "figures", "jobs-percent-yes.png")
ggsave(filename = figure_file, plot=p14) 


p13
p14
```
On the first figure, we can see that jobs with the most positive outcomes are management, technician, and blue-collar, but on the second chart we can see that the job-types with the
highest percentages are 'student', 'retired', and 'unemployed.' There appears to be a noticable different between the total number of positive cases, and the job types which are most commonly recording positive cases.



```{r}
data_location <- here::here("data","processed-data","processeddata5.rds")
mydata <- readRDS(data_location)

numericonly = mydata %>% 
  select(!y_termSubscribed)


# Specify the file path where the image will be saved
file_path <- here("results", "figures","corplot.png")

# Open a PNG graphics device
png(filename = file_path, width = 800, height = 600)

# Create the correlation plot
corrplot(cor(numericonly))

# Close the graphics device
dev.off()
```



